---
title: "Multilanguage annotation with UDpipe"
subtitle: "Parsing time"
author: "Claude Grasland"
output: html_notebook
---

## Introduction

The aim of this short note is to test the time of parsing



## Install the package

Contrary to Spacyr, phe R package UDpipe does not implies a connection with Python and can work directly after installation.

```{r}
#install.package(udpipe)
library(udpipe)
library(knitr)
library(data.table)
```


## Load dataset


Our test data sample is a set of 781131 news published by four daily newspaper  mid 2013 and mid 2020 where at least one country or one world region has been detected by a preliminary research.


```{r}
test<-readRDS("UDpipe/dt_fr_geo_test.RDS")
kable(head(test))

```

## fusion of title and description

We create a column of text called "news" where we add Title and description

```{r}
test$news <- paste(test$title, test$text, sep=". ")
kable(test$news[1])
head(test$news)
```



## Parse text with universal dependencies



## Load french parsing model

we load the default french parsing model

```{r}
sel<-test[test$lang=="fr",]
dl <- udpipe_download_model(language="french")
ud_model_fr<-udpipe_load_model(file=dl$file)
```

## Parse

## Test on 2013-2014 (132201 news)


```{r}
sel<-test[test$date<as.Date("2015-01-01"),]
t1<-Sys.time()
k<-dim(test)[1]
sel<-test[1:k,]
meta<-data.frame(doc_id=sel$id,source=sel$source,date=sel$date,tags=sel$tags)
res<-udpipe_annotate(ud_model_fr, x=sel$news,doc_id=sel$id )
res<-as.data.frame(res)[,c(1,3,5:8,10:12)]
res<-merge(meta,res,all.x=F,all.y=T,by="doc_id")
t2<-Sys.time()

paste("Job done in ",t2-t1)


saveRDS(res,"UDpipe/France_2013_2014_parsed")
```

