---
title: "UDpipe test"
output: html_notebook
---

```{r}
library(udpipe)
library(dplyr)
library(data.table)
library(knitr)

```


# Prepare data

We have prepared a sample of news published in 2019 by different newspapers. We have randomly selected for each newspapers a sample of 1000 news containing at least a country (name or capital city) or a world region (based on our dictionary). We start by working on the french corpus of 1000 news from Figaro and 1000 news from Le Monde. 

```{r}
corpus<-readRDS("UDpipe/samp2019.RDS")
corpus$lang<-substr(corpus$source,1,2)
corpus$num<-row.names(corpus)
corpus$id<-paste(corpus$source,corpus$date,corpus$num,sep="_")
corpus_fr<-corpus %>% filter(lang=="fr")
```



# Annotate

## Download the french language model

```{r}
dl <- udpipe_download_model(language="french")
str(dl)
```

## Load the french model for use

```{r}
ud_model_fr<-udpipe_load_model(file=dl$file)
```

## Annotate the text


### French Trial

**Warning** : it is important to convert the text in "UTF-8" before to run the model. It can produce errors if it is not the case.


```{r}
txt<-c("L'Europe ne se fera pas d'un coup.", 
       "La Russie menace l'Europe de représailles.",
       "L'Europe est menacée par la Russie.",
       "Le conflit se passe en Europe.",
       "Il faut se battre pour l'Europe.",
       "Macron déclare que c'est une victoire de l'Europe.")
txt<-iconv(txt, to="UTF-8")
res<-udpipe_annotate(ud_model_fr, x=txt )

y<-as.data.table(res)
y[,-c(1:4,9:10,13:14)]
```

### French corpus

We check the time necessary fo the processing of 2000 sentences (titles of news)

```{r}
txt<-corpus_fr$text
t1<-Sys.time()

#txt<-iconv(txt, to="UTF-8")
res<-udpipe_annotate(ud_model_fr, x=txt,doc_id =corpus_fr$id)
t2<-Sys.time()
paste("Job done in",t2-t1)
y<-as.data.table(res)
y<-y[,-c(2:4,9:10,13:14)]


```

The time is equal to 13 seconds which mean approx. 6 to 7 seconds / 1000 news or one minute for 10000 news which is reasonable. 




